---
output: github_document
---

## **FunSurv**

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

### Funcational data analysis for recurrent event data
The [***FunSurv*** package](https://htmlpreview.github.io/?https://raw.githubusercontent.com/zifangkong/FunSurv/refs/heads/master/docs/index.html) provides a novel framework for the joint analysis of recurrent event and longitudinal data using a functional regression approach. This package applies functional data analysis techniques to model the time-varying effects of longitudinal predictors on recurrent event processes, addressing the complex dependence structure between the two data types. 

### Installation

You can install and load **FunSurv** from CRAN using

```{r cran-installation, eval = FALSE}
install.packages("FunSurv")
library(FunSurv)
```

You can install **FunSurv** from GitHub with:

```{r gh-installation, eval = FALSE}
## install.packages("devtools")
devtools::install_github("zifangkong/FunSurv", ref = "master")
```

### Citation
Cite ***FunSurv*** with `citation("FunSurv")`.
```{R, eval = FALSE}
citation("FunSurv")
```

### Online documentation 
Online documentation is available [here](https://htmlpreview.github.io/?https://raw.githubusercontent.com/zifangkong/FunSurv/refs/heads/master/docs/articles/funsurv-AR1_FRAILTY.html).

### References:
   Gellar, J. E., Colantuoni, E., Needham, D. M., & Crainiceanu, C. M. (2015). Cox Regression Models with Functional Covariates for Survival Data. *Statistical Modelling*, **15**(3), 256–278. https://doi.org/10.1177/1471082X14565526

  [UNDER REVIEW] Kong, Z., Chiou, S.H., Maalouf, N.M., & Liu Y.L. (2025+) Joint analysis of longitudinal and recurrent event data: A functional regression approach with autoregressive frailty. 
  
  Kong, D., Ibrahim, J. G., Lee, E., & Zhu, H. (2018). FLCRM: Functional linear cox regression model. *Biometrics*, **74**(1), 109-117.
  
  Yao, F., Müller, H. G., & Wang, J. L. (2005). Functional linear regression analysis for longitudinal data. *The Annals of Statistics*, **33**(6), 2873–2903.
  
  Yao, Fang. Functional principal component analysis for longitudinal and survival data. *Statistica Sinica*, (2007): 965-983.
 
  Yau, K. K. W., & McGilchrist, C. A. (1998). ML and REML estimation in survival analysis with time dependent correlated frailty. *Statistics in Medicine*, **17**(11), 1201-1213.
  

  