---
title: "Joint analysis of longitudinal and recurrent event data: A functional regression approach with autoregressive frailty"
author: "Zifang Kong"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    toc: true
    toc_depth: 5
    toc_float: true
bibliography: ../inst/bib/mybib.bib
vignette: >
  %\VignetteIndexEntry{Joint analysis of longitudinal and recurrent event data: A functional regression approach with autoregressive frailty}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{css, echo=FALSE}
pre {
  white-space: pre-wrap;
  word-wrap: break-word;
}

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  options(width = 100)
)
```

In this vignette, we aim to demonstrate how to use our package to jointly analyze longitudinal and recurrent event data using functional regression approach with autoregressive frailty. 

```{r load-pkg, message = FALSE}
library(devtools)
load_all()
library(MFPCA)
library(FunSurv)
packageVersion("FunSurv")
```

## Import data

Suppose we have collected data on both longitudinal measurements and recurrent event outcomes. The **surv_data** dataset contains information on recurrent event outcomes, including the event start time (*t_start*), event end time (*t_stop*), censoring time (*censoring_time*), event status (*status*), and a scalar covariate (*z1*).

```{r surv_data}
load("../data/surv_data.RData")
head(surv_data)
```
The **fun_data** contains longitudinal measurements (*x*), recorded at multiple time points (*time*), for each subject (*id*) 
```{r fun_data}
load("../data/fun_data.RData")
head(fun_data)
```

We use the *plotEvents* function in **reReg** package (@chiou2023regression) to create a recurrent event plot. The horizontal axis represents the subjects' follow-up time, while the vertical axis represents individuals, sorted by the follow-up duration. Each green dot indicates a recurrent event. 

```{r plot_events, fig.width=6, fig.height=4}
library(reReg)
reReg::plotEvents(with(surv_data, Recur(t_stop, id, status)), xlab="Follow-up time", ylab="Subjects", cex=1,recurrent.color="darkgreen", control=list(alpha = 0.8, base_size=8))
```

## Fit the AR1_FRIALTY model

We used a two-step procedure to fit the joint model. 

### Step 1

First, we apply the functional principal component analysis (FPCA) for the longitudinal outcome and obtain the event-specific FPC scores. The calculation of event-specific FPC scores involves 1) calculate the overall FPC scores using *PACE* function in the **MFPCA** package (@mfpca); 2) incorporate the overall FPC scores to obtain the event-specific FPC scores.
```{r PACE_args}
args(AR1_PACE)
```

* `fun_data` is data frame containing longitudinal measurements.
* `surv_data` is a data frame containing information of recurrent event outcomes.
* `nbasis` is an integer specifying the number of B-spline basis functions used for
estimating the mean function and performing bivariate smoothing of the covariance surface (see *fpca.sc* in the **refund** package @crainiceanu2013package).
* `pve` is numerical value between 0 and 1, representing the proportion of variance explained used to choose the number of functional principal components. Defaults to 0.9.
* `npc` is an integer specifying the number of functional principal components.
* `makePD`is boolean value for should positive definiteness be enforced for the covariance surface estimate? (cf. *fpca.sc* in **refund** @crainiceanu2013package).
* `cov.weight.type` defines the type of weighting used for the smooth covariance estimate. Defaults to "none", i.e. no weighting. Alternatively, "counts" (corresponds to *fpca.sc* in **refund** @crainiceanu2013package) weights the pointwise estimates of the covariance function by the number of observation points.


```{r PACE}
fpca = AR1_PACE(fun_data, surv_data, nbasis=10, pve=0.9)
head(fpca$window_scores_fpca)
plot(fpca$uni.PACE$functions, xlab = "Follow-up time", lwd = 3, col = c("blue4","violet", "orange1"))
legend("bottomright", col = c("blue4","violet", "orange1"), lty = 1, lwd = 3, bty = "n", cex = 1, legend = c("FPC1", "FPC2", "FPC3"))
```

By fitting the *AR1_PACE* function, we obtain a data frame of 3 columns of FPC scores, where each column corresponds to a leading functional principal component. 


### Step 2
In the second step, we incorporate the FPC scores obtained from the previous step, and 
```{r}
args(AR1_FRAILTY)
```
* `formula` A Event formula, with the response on the left of a ~ operator, and scalar covariates on the right. The response must be a Event object using Event function.
* `data` A data frame, including time-to-event outcomes, and scalar coefficients
* `scores` A data frame of FPC scores, returned by the *AR1_PACE* function
* `para0` Initial values for $\theta^2$ and auto-regressive coefficient $\rho$.
* `iter.max` Maximum number of iterations for both inner iteration and outer iteration.
* `eps` Tolerance criteria for a possible infinite coefficient value.

```{r}
surv_data = surv_data[surv_data$id %in% fun_data$id, ]
model = AR1_FRAILTY(formula=Event(t_start, t_stop, id, status) ~ z1, data=surv_data, fpca_obj=fpca, para0=c(0.5, 0.5))
```


### Plot

The functional coefficient $\beta(t)$ can be plotted via `plot()`:
```{r plot-betat}
plot_beta(model)
```

