---
title: "Joint analysis of longitudinal and recurrent event data: A functional regression approach with autoregressive frailty"
author: "Zifang Kong"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    toc: true
    toc_depth: 5
    toc_float: true
bibliography: ../inst/bib/mybib.bib
vignette: >
  %\VignetteIndexEntry{Joint analysis of longitudinal and recurrent event data: A functional regression approach with autoregressive frailty}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{css, echo=FALSE}
pre {
  white-space: pre-wrap;
  word-wrap: break-word;
}

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  options(width = 100)
)
```

In this vignette, we aim to demonstrate how to use our package to jointly analyze longitudinal and recurrent event data using functional regression approach with autoregressive frailty. 

```{r load-pkg, message = FALSE}
library(devtools)
load_all()
library(FunSurv)
packageVersion("FunSurv")
```
### Model Formulation

Let $T_{ij}$ denote the time of the $j$th event for subject $i$, and let $C_i$ represent the censoring time. The observed event time, accounting for right censoring, is $\tilde{T}_{ij}=\min(T_{ij}, C_i)$, and $\delta_{ij}=I(T_{ij}\leq C_i)$ serves as an indicator of whether the $j$th event for subject $i$ is observed.
Building on the survival analysis framework with time-dependent, correlated frailty introduced by @yau1998ml, the hazard function is specified as
\begin{equation}\label{hazardFunction}
    \begin{aligned}
h(t; \boldsymbol{Z}_i, {X}_i(\cdot))=h_{0}(t-t_{i,j-1}) \exp \left(\eta_{ij}\right), 
    \end{aligned}
\end{equation}
where $h_0(\cdot)$ is the baseline hazard function, and $\eta_{ij} = \boldsymbol{\alpha}^{\top}\boldsymbol{Z}_i +\int_{t_{i, j-1}}^{t}{X}_{i}(s)\beta(s)ds + v_{ij}$. Here, $t_{i, j-1}$ is the previous event time with $t_{i0} = 0$. $\boldsymbol{\alpha}$ is the fixed effect parameter associated with the time-invariant covariates $\boldsymbol{Z}_i$, and $\beta(t)$ is a time-varying coefficient that captures the effect of functional predictor ${X}_{i}(t)$ on the hazard rate of recurrent events. 



### Import data

Suppose we have collected data on both longitudinal measurements and recurrent event outcomes. The **sdat** dataset contains information on recurrent event outcomes, including the event start time (*t_start*), event end time (*t_stop*), censoring time (*censoring_time*), event status (*status*), and a scalar covariate (*z1*).

```{r sdat}
data("simDat")
head(sdat)
```
The **fdat** contains longitudinal measurements (*x*), recorded at multiple time points (*time*), for each subject (*id*) 
```{r fdat}
head(fdat)
```

We use the *plotEvents* function in **reReg** package (@chiou2023regression) to create a recurrent event plot. The horizontal axis represents the subjects' follow-up time, while the vertical axis represents individuals, sorted by the follow-up duration. Each green dot indicates a recurrent event. 

```{r plot_events, fig.width=6, fig.height=4}
reReg::plotEvents(with(sdat, Recur(t_stop, id, status)), xlab="Follow-up time", ylab="Subjects", cex=1,recurrent.color="darkgreen", control=list(alpha = 0.8, base_size=8))
```

### Fit the AR1_FRIALTY model

The *AR1_FRIALTY* model employs a two-step procedure to fit the joint model.
First, we apply the functional principal component analysis (FPCA) for the longitudinal outcome and obtain the event-specific FPC scores. The calculation of event-specific FPC scores involves 1) calculate the overall FPC scores using *PACE* function in the **MFPCA** package (@mfpca); 2) incorporate the overall FPC scores to obtain the event-specific FPC scores. The results of the first step is a data frame of 3 columns of FPC scores, where each column is associated with a leading functional principal component. 

In the second step, the obtained scores are incorporated into a dynamic recurrent frailty model with an autoregressive structure to account for within-subject correlations across recurrent events.

```{r}
args(AR1_FRAILTY)
```
* `formula` A formula, with the response on the left of a ~ operator being a *Recur* object as returned by function Recur in **reda**, and scalar covariates on the right.
* `sdat` is a data frame containing information of recurrent event outcomes.
* `fdat` is data frame containing longitudinal measurements.
* `para0` A vector of initial values for $\theta^2$ and auto-regressive coefficient $\rho$. Both defaults to 0.5.
* `nbasis` is an integer specifying the number of B-spline basis functions used for
estimating the mean function and performing bivariate smoothing of the covariance surface (see *fpca.sc* in the **refund** package @crainiceanu2013package).
* `pve` is numerical value between 0 and 1, representing the proportion of variance explained used to choose the number of functional principal components. Defaults to 0.9.
* `npc` is an integer specifying the number of functional principal components.
* `makePD`is boolean value for should positive definiteness be enforced for the covariance surface estimate? (cf. *fpca.sc* in **refund** @crainiceanu2013package).
* `cov.weight.type` defines the type of weighting used for the smooth covariance estimate. Defaults to "none", i.e. no weighting. Alternatively, "counts" (corresponds to *fpca.sc* in **refund** @crainiceanu2013package) weights the pointwise estimates of the covariance function by the number of observation points.
* `iter.max` Maximum number of iterations for both inner iteration and outer iteration.
* `eps` Tolerance criteria for a possible infinite coefficient value.

We set the percentage of variance explained to 90\% to determine a finite number of FPC components, resulting in three FPC components
```{r}
fit <- AR1_FRAILTY(formula=Recur(t_start %to% t_stop, id, status) ~ z1, sdat=sdat, fdat=fdat, nbasis=10, pve=0.9)
summary(fit)
```

The functional coefficient $\beta(t)$ can be plotted via `plot()`:
```{r plot-betat}
plot(fit, which="beta")
```

The plot below shows the estimated baseline survival probability. 
```{r plot-basesurv}
plot(fit, which="basesurv")
```

The plot below demonstrates the three functional principal components estimated from longitudinal measurements.
```{r plot-fpc}
plot(fit, which="fpc")
```


### References
